### 2016-06-11 aplas16/main-aplas.tex
Created a directory for aplas16.

\documentclass{llncs}

\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amsthm}
\usepackage{url}
\usepackage{mathptmx}
\usepackage{doubleequals}
\usepackage{makeidx}  % allows for indexgeneration

\input{Preamble}
\input{macro}

\begin{document}

\title{Preventing Runtime Errors of Redis at Compile Time}
\author{Ting-Yan Lai \inst{1}\inst{2}\and Tyng-Ruey Chuang \inst{1}
  \and Shin-Cheng Mu \inst{1}}
\institute{Institute of Information Science, Academia Sinica, Taiwan
\and Institute of Computer Science and Engineering,
National Chiao Tung University, Taiwan}

\maketitle

\begin{abstract}
Programmers often interact with database systems by sending queries through libraries or packages in some programming languages. People, however, make mistakes.
While some of the syntactic and semantic errors can be prevented by the
 language at compile time, or caught by the package at runtime, most semantic
 errors are just being ignored, causing problems at the database system.

In this paper, we demonstrate how to prevent those runtime errors at compile
 time, thus allowing users to write more reliable database queries without
 runtime overhead, by exploiting type-level programming techniques such as
 indexed monad, type-level literals and closed type families in Haskell.

The database system and the package we are targeting are \emph{Redis} and
 \emph{Hedis} respectively, and our implementation is available as \emph{Popcorn}
 on \emph{Hackage}.
\end{abstract}

\input{sections/Introduction}
\input{sections/Motivation}
\input{sections/TypeLevelDict}
\input{sections/IndexedMonads}
\input{sections/TypeLevelFun}
\input{sections/PolyRedis}
\input{sections/Constraints}
\input{sections/Assertions}
\input{sections/Discussions}
\input{sections/Related}
\input{sections/Conclusions}

\bibliographystyle{splncs03}
\bibliography{cites}

\end{document}