### 2016-06-11 aplas16/sections/Related.tex

\section{Related Work}
\label{sec:related}

While Redis can be viewed as a non-relational database system (although Redis
 seems reluctant to admit this in recent years), there also are similar goals on
 relational database systems, trying to achieve a safer database interface by
 making them statically checked.
 \emph{HaskellDB}\cite{haskelldb, haskelldbimproved} expresses quires
 with relational algebra-like combinators, wrapped in phantom types. Or as
 examples, at first to demonstrate the power of dependent types in
 Agda\cite{pi}, and then with singletons in Haskell\cite{singletons}.