### 2016-06-16 aplas16/sections/Conclusions.tex
A very brief Conclusion.

\section{Conclusions}
\label{sec:conclusions}

By exploiting various recent extensions and type-level programming techniques,
we have designed a domain-specific embedded language \Edis{} which, at
term-level, is simply a wrapper of \Hedis{}, but enforces typing disciplines that keeps track of available keys and their types and allows a program to be
constructed only if it does not throw a runtime type error. We believe that it
is a neat case study of application of type-level programming.

% With more and more extensions added to the language, Haskell is gradually
%  becoming a dependently-typed language,\footnote{Why not be dependently typed? \url{http://stackoverflow.com/questions/12961651/why-not-be-dependently-typed}}
%  but it's not that dreadful as many (including us) would have thought.