### 2016-06-17 aplas16/main-aplas.tex
Another round of revision from the beginning. Corrected some bibliography.

\documentclass{llncs}

\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amsthm}
\usepackage{color}
\usepackage{url}
\usepackage{mathptmx}
\usepackage{doubleequals}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{textcomp} % for single quote
\usepackage{fancyvrb} % for indented verbatim

\input{Preamble}
\input{macro}

\begin{document}

\title{Type Safe Redis Queries}
\subtitle{A Case Study of Type-Level Programming in Haskell}
\author{Ting-Yan Lai \inst{1}\inst{2}\and Tyng-Ruey Chuang \inst{1}
  \and Shin-Cheng Mu \inst{1}}
\institute{Institute of Information Science, Academia Sinica, Taiwan
\and Institute of Computer Science and Engineering,
National Chiao Tung University, Taiwan}

\maketitle

\begin{abstract}
\Redis{} is an in-memory data structure store, often used as a database, with a
Haskell interface \Hedis{}. \Redis{} is dynamically typed --- a key can be
discarded and re-associated to a value of a different type, and a command,
when fetching a value of a type it does not expect, signals a runtime error. We
develop a domain-specific language that, by exploiting Haskell type-level
programming techniques including indexed monad, type-level literals and closed
type families, keeps track of types of values in the database and statically
guarantees that type errors cannot happen for a class of \Redis{} programs.
\end{abstract}

\input{sections/Introduction}
\input{sections/IndexedMonads}
\input{sections/TypeLevelDict}
\input{sections/Embedding}
\input{sections/Discussions}
\input{sections/Conclusions}

\bibliographystyle{splncs03}
\bibliography{cites}

\end{document}